---
title: "Import explore data"
author: "Kundan K. Rao"
date: "01/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Basic data cleaning and exploration

loading dataset in the workspace

# importing the wine dataset

```{r}
library(readr)
data <- read_csv("Wine 0.csv - Wine 0.csv.csv", show_col_types = F)
```

# Check structure of the dataset and data types of variables
```{r}
# structure
str(data)
```
Observation : 
  
  1.   Every variable is of numeric (double) data types
  2.   There are 178 observation of 13 variables



# Head and tail of the data set

```{r}
# head
head(data)
```

```{r}
# tail
tail(data)
```

# Checking for null and other types of missing data

```{r}
# any NA in the dataframe
any(is.na(data))
```

```{r}
# total NAs
sum(is.na(data))
```

```{r}
# checkinf if there is any infinite valuesm
f<-function(x){any(is.infinite(x))}
sapply(data,f)
```

```{r}
# Checking for total infinite values
f <- function(x){sum(is.infinite(x))}
sapply(data,f)
```

```{r}
# checking for NaN values
f<-function(x){any(is.nan(x))}
sapply(data,f)
```

```{r}
# total NaN values
f <- function(x){sum(is.nan(x))}
sapply(data,f)
```

Observation: There are neither infinity nor ant NaN values in any variable in whole dataset.
